---
title: "K-Means-Iris"
author: "Mandy"
date: "2015.11.3"

## Does k-means work well for classifying this dataset? Does scaling help?
```{r,echo=F}
data("iris")
head(iris)
irisnew <- iris[-5]
```

```{r}
# choosing number of clusters
require("NbClust")
nc <- NbClust(irisnew, min.nc=2, max.nc=15, method="kmeans")
barplot(table(nc$Best.n[1,]),xlab="Num of Clusters", ylab="Num of Criteria")
```

We could see from the plot that 2 clusters are the best. This also agrees with the barplot results with 2 clusters being supported the most by 10 indicies. However, 3 clusters are the second suggested one with 8 proposed indices.  
```{r}
wss <- 0
for (i in 1:15){
  wss[i] <-sum(kmeans(irisnew, centers=i)$withinss)
}
plot(1:15,wss,type="b", xlab="Number of Clusters", ylab="Within groups sum of squares")
```

The sum of square error (SSE) screen plot also shows that cluster=2 is the significant bend. Both methods suggest k=2 is the best, but as there are 3 types in iris, we continue to try 2 and 3 clusters respectively.

```{r}
fit.kmiris2 <- kmeans(irisnew,2)
fit.kmiris2
fit.kmiris3 <- kmeans(irisnew,3)
fit.kmiris3
library(fpc)
plotcluster(irisnew, fit.kmiris2$cluster)
plotcluster(irisnew, fit.kmiris3$cluster)
```

when cluster=2, (between_SS / total_SS =  77.6 %) thus the overall difference is very large between groups. Also, there are almost no collapse between clusters, but there are 3 data points in group 2 very close to group 1.  
When cluster=3, (between_SS / total_SS =  88.4 %) yields a better difference across groups. There is some collapse between group1 and 3 but overall it makes more sense because it's more like the real case.  
Also, it's noticeable that the two clustering numbers give very different results.

```{r}
a=table(iris$Species,fit.kmiris2$cluster)
b=table(iris$Species,fit.kmiris3$cluster)
a
b
```

```{r}
library(flexclust)
randIndex(a)
randIndex(b)
```
A small ARI indicates a poor classification.  
When cluster=2, ARI is only 0.54, which is conceivable because there are only 2 clusters compared to the original 3 groups.  
When cluster=3, ARI is 0.73, which is still not satisfatory so I continue to scale the data.

```{r}
nc <- NbClust(scale(iris[-5]), min.nc=2, max.nc=15, method="kmeans")
barplot(table(nc$Best.n[1,]),xlab="Num of Clusters", ylab="Num of Criteria")
```

After scaling, the two methods all suggest cluster number to be 3.  
```{r}
fit2=kmeans(scale(iris[-5]),3)
plotcluster(scale(iris[-5]), fit2$cluster)
```  
The plot looks like the unscaled plot with cluster=3, however the boundary is even more blur.  
```{r}
c=table(iris$Species,fit2$cluster)
library(flexclust)
randIndex(c)
```  
After scaling, ARI drops to 0.62. K-means seems not a good classfication method for iris and the scaling doesn't help. It is because the predictor in iris are similar in terms of sacle. 

