---
title: "Kmeans_part one"
author: "Tianyu Zhan"
date: "November 3, 2015"
output: html_document
---
  
# K-means  
Use k-means method to cluster the wines into 3 groups.
```{r}
data(wine,package="rattle")
fit=kmeans(wine[-1],3)
# the default measure to calculate k-means is Euclidean distance
library('fpc')
plotcluster(wine[-1],fit$cluster)
```  

From the graph we can see that the boundaries are not very clear, which indicates that some observations can be classified into either groups. 
Draw parallel coorinates plot to see how variables contributed in each cluster.  
```{r}
library(MASS)
parcoord(wine[-1],fit$cluster)
```  
We cannot observe a clear pattern from this graph.  
Take a look at the misclassification rate (number of misclassified obs divided by total number of obs).
```{r,echo=F}
mis1=table(wine$Type,fit$cluster)
mis1
```   
There are 145(=46+20+29+50) out of 178 obs that have been misclassified, so the misclassification rate is 0.81.  
To take a further look at how well the kmeans algorithm did, we use ARI (Adjusted Rand Index) as an evaluation criteria. 
$$ARI=\frac{\Sigma_{ij}\binom{n_{ij}}{2}-[\Sigma_{i}\binom{a_i}{2}\Sigma_{j}\binom{b_j}{2}]/\binom{n}{2}}{\frac{1}{2}[\Sigma_{i}\binom{a_i}{2}+\Sigma_{j}\binom{b_j}{2}]-[\Sigma_{i}\binom{a_i}{2}\Sigma_{j}\binom{b_j}{2}]/\binom{n}{2}}$$
where $n_{ij}$ denotes for the ith row and jth column in a contigency table; $a_i$ denotes for the sum of ith row and $b_j$ denotes for the sum of jth column.
```{r,echo=F}
library(flexclust)
randIndex(mis1)
```   
A small ARI indicates a poor classification.  

We want to improve the performance by __scaling and centering__ (scale() function can do this work) the data.
```{r}
fit2=kmeans(scale(wine[-1]),3)
plotcluster(scale(wine[-1]),fit2$cluster)
parcoord(scale(wine[-1]),fit2$cluster)
mis2=table(wine$Type,fit2$cluster)
mis2
randIndex(mis2)
```  
We can see the data is clustered very well, with no collapse between clusters.  
From the graph, we may consider that black cluster contains wine with low flavanoids value, low proanthocyanins value, low hue value; or green cluster contains wine which has dilution value higher than wine in red cluster.  
This time, the misclassification rate is 0.034, much better than the previous result. ARI is 0.897, also increased a lot.    
Therefore, scaling the data can improve the clustering quality.